### 2011-11-04 doc/report/dispostion2.tex

\documentclass{article} 
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

% ska gå att förstå och locka till läsning
%ska tydligt beskriva vad rapporten handlar om
%ska innehålla vettiga sökord
\title{Dispositon for master thesis \\
      \small{A large scale grammar and parser for Swedish}}
      %First results of developing a Swedish parser based on a wide-coverage grammar\\
        %Combining language resources into a grammar-driven Swedish parser} 
        %or Bootstrapping a Swedish parser \\
        %or A grammar-based Swedish parser from already existing resources}
\author{Malin Ahlberg
\institute{Department of Computer Science \& Engineering, Gothenburg University, Sweden}
}
\begin{document}
\maketitle

\section{Introduction} - 
Short introduction to parsing, GF, swedish, controlled natural language vs.
free language.
Why we want a parser, what we can get from GF and what we can't get.
Which parts we are focusing on and why.

\section{Background} - 
Longer introduction to GF, Talbanken and Saldo. Other parsers and related work.

\section{Progress} - 
One part about how we have imported saldo, problems and results, how we deal
with words that were not imported. Use of lexical aquisition tools.

One part about the mapping from Talbanken to GF, description. This part explains
the tags in Talbanken and gives examples of how they are translated.

One part about the grammar. What parts that I have been working on and why, 
what constructions I wanted to add, examples of sentences that should work
and of sentences that shouldn't be covered. How to separate Swedish from
Scandinavian, bigger reconstructions in the grammar.

\section{Results and discussion}
Evaluation methods and results from all parts.

Discussion of the results and methods, why are the results good, why are they not better.
What other methods could have been used? What did we/I expect and what happend?

\section{Future Work}
what is still left to do? Description and direction of future work:
how to make the parser robust, what is still needed in the grammar, 
how to get valency information, make the parser faster.
Getting and using probabilities.

\section{Conclusion}
My conclusions about the work.