### 2011-11-21 doc/report/Refl.tex
report, and planning

varandra and sin/sitt/sig. Not subject, but NP.
Want the following:
\enumsentence{Han såg sina barns skor.}
\enumsentence{Sina vantar hade han glömt på tåget.}
but not
\enumsentence{*Sina vantar var kvar på tåget.}
\enumsentence{Han ger sina pengar till sina syskon.}
\enumsentence{*Jag ger sina pengar till sina syskon.}
\enumsentence{Han är längre än sin kompis.}
\enumsentence{Hon tyckte om alla sina elever.}
\enumsentence{Han såg sin hund och sitt hus}
\label{ex:refGen}
\enumsentence{Han är längre än sin kompis}
\label{ex:ReflAP}

Suggest a connection between the subject and the object.
Could be combined with 'sig', a ReflNP without another noun.
Make a distinction between objects and subjects?
do not want to copy all rules for another category. Would like depent types!
work pretty well by letting all NP depend on the subject in the phrase.
The current version is nice and works for recursive genitivs!
Extra info, but maybe this is just the case about NPs.
Discuss how other solutions, NPR and Predet : Predet -> CN -> NPR ;
*Slash* : VPSlash -> NPR -> VP ;
Keep number of fields down, but introduces another category, and one more
label to the parse tree, is it interesting? How to combine it with genitive 
without making it ambiguous (Han såg barnens hund).
What if we add alot of other rules that uses NP, will they all need to be
duplicated then?

'han såg sin mammas bil', they same could be applied to
APs for \ref{ex:ReflAP} or Adv for
\enumsentence{Han leker oftare än sin kompis}
 (ComparAdvAdj)
Atm you can't focus this kind of object, then a ClSlash need info about the
subject. FunRP also tricky (and what does it mean, we can say
Han såg sin katt i vilken musen låg), in this case RP needs to give more information.
Kort sagt, the information about the subject must be spread
in different parts of the sentence. 
Could also decide to always produc 'sin' -> jag såg sin bil
to avoid the dependence, but not nice. Or keep a flag in NP saying if it is
in third person, and add nonexist for others. But this way we can also get rid
of ReflVP and the possessive prons \verb|PossPron| should only be used for somebody else:
'jag såg din bil'.
Unfortunaly, have to change ReflVP, this only allows this in object for VPSlash, not FocObj,
not 'alla sina', usw. Want it to handle conjuction as in \ref{ex:refGen}
etc.